---
title: "Princeton University - Panel Data Using R"
author: "Pooya Shirazi"
date: "March 19, 2018"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

# Getting Started in Fixed / Random Effects Models using R

## Introduoction

`Panel Data` (also known as `Longitudinal` or `Cross-sectional Time-series` data) is a dataset in which the behavior of entities are observed across time.

These entities could be states, companies, individuals, countries, etc.

## Exploring Panel Data

```{r eval = TRUE, echo = TRUE, results = "hide", message = FALSE, warning = FALSE}
# Load Required Packages: 
library(tidyverse)
library(plm)
```

```{r}
# Read Data Examples from Local System:
Panel.Data <- read.csv(file = "C:/Users/Pooya/OneDrive/My Tutorials/Academic/Princeton University - Panel Data Using R/Data.csv")

# Compactly Display the Structure of Panel.Data:
str(Panel.Data)
```

```{r}
Panel.Data %>% 
  ggplot(mapping = aes(x = year, y = y, color = country)) +
  geom_point(size = 3) +
  geom_smooth(se = FALSE, method = "loess") + 
  theme(panel.background = element_rect(fill = "white", colour = "black", linetype = 1),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5), aspect.ratio = 1) +
  scale_x_continuous(limits=c(1990,1999), breaks = c(1990:1999)) +
  scale_y_continuous(labels = scales::comma)
```


```{r}
# Exploring Panel.Data:
Panel.Data %>% 
  ggplot(mapping = aes(x = year, y = y)) +
  geom_point() +
  geom_line() + 
  facet_wrap(facets = ~ country, nrow = 2) + 
  theme(panel.background = element_rect(fill = "white", colour = "black"),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5), aspect.ratio = 1) +
  scale_x_continuous(limits=c(1990,1999), breaks = c(1990:1999))
```

## Fixed-effects Model: Heterogeneity Across Countries (or Entities)

```{r}
Panel.Data %>% 
  ggplot(mapping = aes(x=country, y=y)) + 
  stat_summary() +
  theme(panel.background = element_rect(fill = "white", colour = "black"),
        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust = 0.5), aspect.ratio = 1,
        plot.title = element_text(hjust = 0.5)) +
  ggtitle(label = "Heterogeineity Across Countries")

```

## Fixed Effects: Heterogeneity Across Years

```{r}
Panel.Data %>% 
  ggplot(mapping = aes(x=year, y=y)) + 
  stat_summary() +
  theme(panel.background = element_rect(fill = "white", colour = "black"),
        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust = 0.5), aspect.ratio = 1/3,
        plot.title = element_text(hjust = 0.5)) +
  scale_x_continuous(limits=c(1990,1999), breaks = c(1990:1999)) +
  ggtitle(label = "Heterogeineity Across Years")

```

```{r}
library(gplots)
plotmeans(y ~ year, main="Heterogeineity across years", data=Panel.Data)
```

